---
title: 'COVID-19: nowcast and forecast'
author: "Paul Birrell, Joshua Blake, Edwin van Leeuwen, MRC Biostatistics Unit COVID-19 Working Group, Daniela De Angelis."
date: "`r lubridate::today()`"
output:
  html_document:
    self_contained: no
---

<base target="_parent">

```{r preamble, include=FALSE}
if (!exists("external")) external = T
knitr::opts_chunk$set(echo = FALSE, cache.path = "~/tmp/cache")

suppressMessages(library(knitr))
suppressMessages(library(DT))
suppressMessages(library(lubridate))
suppressMessages(library(plotly))
suppressMessages(library(tidyverse))
if(exists("Rfile.loc")) {
  source(file.path(Rfile.loc, "results_api.R"))
  source(file.path(Rfile.loc, "plot_funcs.R"))
} else {
  proj.dir <- "~/RTM"
  source("results_api.R")
  source("plot_funcs.R")
}
results_hash = tools::md5sum(file.path(out.dir, "output_matrices.RData"))
if (external && exists("dth.dat")) {
	dth.dat <- filter(dth.dat, date <= ymd(date.data) - reporting.delay)
	results_hash = NULL
}

```

## Epidemic summary {.tabset}

### Current $R_t$

Value of $R_t$, the average number of secondary infections due to a typical infection today.

```{r Rt-table}
get.aggregated.quantiles(Rt, "region", QUANTILES) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  select(Region = region, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) %>%
  datatable() %>%
  formatRound(2:4)

```

### Number of infections

```{r cum-inf-sum}
format_num <- function(x) {
  return(format(round(signif(x, 3)), big.mark = ",", scientific = FALSE, width = 0))
}

tbl_cum_inf_region <-
  get.aggregated.quantiles(cum_infections, "region", c(0.025, 0.5, 0.975)) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data))
  
tbl_ar_Eng <- get.aggregated.quantiles(cum_infections, NULL, c(0.025, 0.5, 0.975))  %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  mutate(
    population = sum(population$value),
    region = "Total"
  )

tbl_inf_region <-
  bind_rows(
    get.aggregated.quantiles(infections, NULL, c(0.025, 0.5, 0.975))  %>%
    pivot_wider(names_from = quantile) %>%
    filter(date == ymd(date.data)) %>%
    mutate(
      region = "Total"
    ),
    get.aggregated.quantiles(infections, "region", c(0.025, 0.5, 0.975)) %>%
    pivot_wider(names_from = quantile) %>%
    filter(date == ymd(date.data))
  )%>%
  mutate_at(vars(`0.025`, `0.5`, `0.975`), format_num) %>%
  transmute(
    Region = region,
    Daily_infections = paste0(`0.5`, " (", `0.025`, "--", `0.975`, ")")
  )

```

```{r cum-inf-disp}
bind_rows(
  select(tbl_ar_Eng, -population),
  tbl_cum_inf_region
) %>%
  mutate_at(vars(`0.025`, `0.5`, `0.975`), format_num) %>%
  transmute(
    Region = region,
    Cumulative_infections = paste0(`0.5`, " (", `0.025`, "--", `0.975`, ")")
  ) %>%
  left_join(tbl_inf_region, by = "Region") %>%
  datatable()
```

### Attack rate {.tabset}

The percentage of a given group that has been infected.

#### By region

```{r attack-region}
bind_rows(
  tbl_ar_Eng,
  population %>%
    group_by(region) %>%
    summarise(population = sum(value)) %>%
    left_join(tbl_cum_inf_region, by = c('region' = 'region'))
  ) %>%
  select(Region = region, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`, population) %>%
  mutate(
    Median = Median / population,
    `95% CrI (lower)` = `95% CrI (lower)` / population,
    `95% CrI (upper)` = `95% CrI (upper)` / population
    ) %>%
  select(-population) %>%
  datatable() %>%
  formatPercentage(2:4)
```


#### By age

```{r attack-age}
tbl_cum_inf_age <-
  get.aggregated.quantiles(cum_infections, "age", c(0.025, 0.5, 0.975)) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  select(age, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) 

population %>%
  group_by(age) %>%
  summarise(population = sum(value)) %>%
  left_join(tbl_cum_inf_age, by = c('age' = 'age')) %>%
  mutate(
    Median = Median / population,
    `95% CrI (lower)` = `95% CrI (lower)` / population,
    `95% CrI (upper)` = `95% CrI (upper)` / population
    ) %>%
  select(-population) %>%
  arrange(match(age, age.labs)) %>%
  datatable() %>%
  formatPercentage(2:4)
```

### IFR


```{r ifr, eval=!is.null(dth.dat)}
today_cum_inf_by_age <- apply(cum_infections[ymd(date.data)-start.date,,,,drop = FALSE], c("age", "iteration"), sum)
merge.youngest.age.groups <- function(mat, num.to.group = 2) {
  if (num.ages <= 1) {return(mat)}
  add.function <- function(x) {
    to.merge <- x[1:num.to.group]
    to.preserve <- x[(num.to.group+1):length(x)]
    return(c(sum(to.merge), to.preserve))
  }
  result <- apply(mat, "iteration", add.function)
  return(result)
}

overall_ifr <- today_cum_inf_by_age %>%
  merge.youngest.age.groups() %>%
  t() %>%
  `*`(params$prop_case_to_hosp[parameter.to.outputs,]) %>%
  rowSums() %>%
  `/`(colSums(today_cum_inf_by_age)) %>%
  quantile(QUANTILES) %>%
  enframe(name = "quantile") %>%
  mutate(age = "Overall")

params$prop_case_to_hosp %>%
  as.matrix() %>%
  `colnames<-`(dimnames(deaths)$age) %>%
  as_tibble() %>%
  pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
  group_by(age) %>%
  summarise(value = list(quantile(ifr, QUANTILES))) %>%
  mutate(quantile = list(names(value[[1]]))) %>%
  unnest(c(quantile, value)) %>%
  bind_rows(overall_ifr) %>%
  mutate(
    quantile = parse.percentage(quantile),
    value = paste0(signif(100 * value, 2), "%")
  ) %>%
  pivot_wider(names_from = quantile, values_from = value) %>%
  select(Age = age, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) %>%
  arrange(match(Age, c("Overall", dimnames(deaths)$age))) %>%
  datatable()

# (params$prop_case_to_hosp %>%
#   as.matrix() %>%
#   `colnames<-`(dimnames(deaths)$age) %>%
#   as_tibble() %>%
#   pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
#   mutate(age = factor(age, levels = dimnames(deaths)$age)) %>%
#   ggplot(aes(x = age, y = ifr)) +
#   geom_violin()) %>%
#   ggplotly()
#   
# params$prop_case_to_hosp %>%
#   as.matrix() %>%
#   `colnames<-`(dimnames(deaths)$age) %>%
#   as_tibble() %>%
#   pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
#   group_by(age) %>%
#   summarise(value = list(quantile(ifr, c(0.025, 0.25, 0.5, 0.75, 0.975)))) %>%
#   mutate(quantile = list(names(value[[1]]))) %>%
#   unnest(c(quantile, value)) %>%
#   mutate(
#     age = factor(age, levels = dimnames(deaths)$age),
#     quantile = parse.percentage(quantile)
#   ) %>%
#   pivot_wider(names_from = quantile, values_from = value, names_prefix = "q") %>%
#   ggplot(aes(x = age, lower=q0.25, upper=q0.75, middle=q0.5, ymin=q0.025, 
#                     ymax=q0.975)) +
#   geom_boxplot(stat = "identity") +
#   scale_y_continuous()
```

### Change in infections incidence {.tabset}

#### Growth rates

```{r growth-rate-func}
signed_inf <- function(num) {
  ifelse(
    num == 0, 0,
    ifelse(num > 0, +Inf, -Inf)
  )
}

add.England <- function(data) {
  sum.all(infections)
  data <- apply(infections, c("region", "iteration", "date"), sum)
  data <- aperm(data, c("date", "region", "iteration", "age"))
  data["Total",,] <- apply(infections, c("iteration", "date"), sum)
}

today.int <- as.integer(ymd(date.data))
wanted.days <- as.character(c(today.int - 1, today.int, today.int + 1))
calc.growth.rate <- function(data) {
  data <- data %>%
    aperm(c("date", "region", "iteration", "age")) %>%
    `[`(wanted.days,,,,drop = FALSE)
  num.iterations <- dim(data)[3]
  data %>%
    apply(c("region", "iteration", "date"), sum) %>%
    as.tbl_cube(met_name = "value") %>%
    as_tibble() %>%
    bind_rows(
      apply(data, c("iteration", "date"), sum) %>%
      as.tbl_cube(met_name = "value") %>%
      as_tibble() %>%
      mutate(region = "Total")
    ) %>%
    mutate(
      date = lubridate::as_date(date)
    ) %>%
    arrange(date, iteration) %>%
    group_by(region, iteration) %>%
    mutate(gradient = lead(value) - lag(value)) %>%
    summarise(growth = 
                ifelse(value == 0,
                       signed_inf(gradient),
                       gradient / (2 * value))[2]
    ) %>%
    group_by(region) %>%
    summarise(Median = median(growth),
              `95% CrI (lower)` = quantile(growth, 0.025),
              `95% CrI (upper)` = quantile(growth, 0.975)
              ) %>%
    arrange(match(region, c("Total", regions))) %>%
    mutate(Region = str_replace_all(region, "_", " ")) %>%
    select(Region, everything()) %>%
    select(-region)
}
infections.growth.rates <- calc.growth.rate(infections)
if (!is.null(deaths)) deaths.growth.rates <- calc.growth.rate(deaths)
```

NB: negative growth rates are rates of decline. Values are daily changes.

`r knitr::kable(infections.growth.rates, digits = 2, align = 'lccc')`

#### Halving times

Halving times in days, if a region shows growth than value will be NA.

```{r inf-halving}

infections.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. > 0, NA, -1/log2(1+.))
  ) %>%
  knitr::kable(digits = 2, align = 'lccc')

```

#### Doubling times

Doubling times in days, if a region shows decline then the value will be NA.

```{r inf-doubling}

infections.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. < 0, NA, log(2)/(.))
  ) %>%
  rename(tmp = `95% CrI (upper)`) %>%
  rename(`95% CrI (lower)` = tmp,
  	 `95% CrI (upper)` = `95% CrI (lower)`) %>%
  select(Region, Median, `95% CrI (lower)`, `95% CrI (upper)`) %>%
  knitr::kable(digits = 2, align = 'lccc')
    
```

### Change in deaths incidence {.tabset}


#### Growth rates

NB: negative growth rates are rates of decline. Values are daily changes.

```{r death-growth, eval=!is.null(deaths)}
knitr::kable(deaths.growth.rates, digits = 2, align = 'lccc')
```

#### Halving times

Halving times in days, if a region shows growth than value will be NA.

```{r death-halving, eval=!is.null(deaths)}

deaths.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. > 0, NA, -1/log2(1+.))
  ) %>%
  knitr::kable(digits = 2, align = 'lccc')

```

#### Doubling times

Doubling times in days, if a region shows decline then the value will be NA.

```{r death-doubling, eval=!is.null(deaths)}

deaths.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. < 0, NA, log(2)/(.))
  ) %>%
  rename(tmp = `95% CrI (upper)`) %>%
  rename(`95% CrI (lower)` = tmp,
  	 `95% CrI (upper)` = `95% CrI (lower)`) %>%
  select(Region, Median, `95% CrI (lower)`, `95% CrI (upper)`) %>%
  knitr::kable(digits = 2, aligh = 'lccc')

```

## Infections and deaths {.tabset}

The blue lines is show when interventions have been introduced (lockdown on 23 Mar and the relaxation of measures on 11 May), and the red line shows the date these results were produced (`r format(ymd(date.data), "%d %b")`).

### Infection incidence {.tabset}

#### By region
```{r reg-inf, cache=TRUE, cache.extra=results_hash}
make.plots(infections, by = "region")
```

#### By age
```{r age-inf, cache=TRUE, cache.extra=results_hash}
make.plots(infections, by = "age")
```

### Cumulative infections {.tabset}

#### By region
```{r reg-cum-inf-reg, cache=TRUE, cache.extra=results_hash}
make.plots(cum_infections, by = "region")
```

#### By age
```{r r-cum-inf-age, cache=TRUE, cache.extra=results_hash}
make.plots(cum_infections, by = "age")
```

```{r eval=!is.null(dth.dat), child='deaths-report.Rmd'}
```


```{r eval=!is.null(hosp.dat), child='hosp-report.Rmd'}
```

### Prob $R_t > 1$

The figure below shows the probability that $R_t$ is greater than 1 (ie: the number of infections is growing) in each region over time.
Clicking the regions in the legend allows lines to be added or removed from the figure.

```{r P-R-gt-1, cache=TRUE, cache.extra=results_hash}
(Rt > 1) %>%
  apply(c("date", "region"), sum) %>%
  `/`(length(dimnames(Rt)$iteration)) %>%
  as.tbl_cube(met_name = "value") %>%
  as_tibble() %>%
  mutate(date = as_date(date)) %>%
  filter(date <= ymd(date.data)) %>%
  create.base.subplot(1, "") %>%
  add_lines(y = ~value, x = ~date, color = ~region) %>%#, hovertemplate = "%{y:.2r}") %>%
  layout(showlegend = TRUE, yaxis = list(tickformat = "%", title = "$\\text{Probability }R_t > 1$"),
         xaxis = list(title = "Date"))
```

### $R_t$
```{r Rt-plot, cache=TRUE, cache.extra=results_hash}
make.plots(Rt, by = "region", y.format = ".2f", combine.to.England = NULL)
```

# 

```{r eval=(!external & file.exists(file.path(out.dir, "projections.RData"))), child='projinf-report.Rmd'}
```

#

Copyright © MRC Biostatistics Unit, University of Cambridge
