---
title: 'COVID-19: nowcast and forecast'
author:
- Paul Birrell
- Joshua Blake
- Edwin van Leeuwen
- Daniela De Angelis
- Joint PHE Modelling Cell
- MRC Biostatistics Unit COVID-19 Working Group
date: "`r lubridate::today()`"
output:
  html_document:
    self_contained: no
---

```{r preamble, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
QUANTILES <- c(0.025, 0.5, 0.975)

suppressMessages(library(knitr))
suppressMessages(library(DT))
suppressMessages(library(lubridate))
suppressMessages(library(plotly))
suppressMessages(library(tidyverse))
if(exists("Rfile.loc")) {
  source(file.path(Rfile.loc, "results_api.R"))
} else {
  proj.dir <- "~/RTM"
  source("results_api.R")
}

```

```{r combine-Rt}
inf_by_region <- apply(infections, c("iteration", "date", "region"), sum)[seq(from = 1, to = length(parameter.to.outputs), length.out = 500),,]
Rt <- aperm(Rt, names(dimnames(inf_by_region)))
stopifnot(all.equal(dim(Rt), dim(inf_by_region), check.names = FALSE))
weighted_Rt_sum <- apply(inf_by_region * Rt, c("iteration", "date"), sum)
inf_by_day <- apply(inf_by_region, c("iteration", "date"), sum)
stopifnot(all.equal(dim(weighted_Rt_sum), dim(inf_by_day)))
overall_Rt <- weighted_Rt_sum / inf_by_day
```


```{r plot-function}
make.plots <- function(projections, ylab = "", by = NULL, data = NULL,
                       y.format = ".3s", combine.to.England = sum.all,
                       combine.data.to.England = sum.all.data) {

  Eng_projection <- combine.to.England(projections) %>% add.quantiles(NULL, QUANTILES)
  Eng_data <- combine.data.to.England(data)
  projections <- get.aggregated.quantiles(projections, by, c(0.025, 0.5, 0.975)) %>%
    rename(by = all_of(by))
  plot.names <- unique(projections$by)
  num.rows <- ceiling(length(plot.names)/2)
  plot.height <- num.rows * 420 + 150
  date <- ymd(date.data)
  lines <- list(
    list(
        type = "line", 
        y0 = 0, 
        y1 = 1, 
        yref = "paper",
        x0 = ymd(date.data), 
        x1 = ymd(date.data), 
        line = list(color = "red"),
        hoverinfo = "Today"
    ),
    list(
        type = "line", 
        y0 = 0, 
        y1 = 1, 
        yref = "paper",
        x0 = ymd(20200323), 
        x1 = ymd(20200323), 
        line = list(color = "blue"),
        hoverinfo = "Lockdown"
    )
  )
  title <- function(name) {
    list(
      text = str_replace_all(name, "_", " "),
      xref = "paper",
      yref = "paper",
      yanchor = "top",
      xanchor = "left",
      x = 0.1,
      y = 0.9,
      font = list(size = 20),
      showarrow = FALSE
    )
  }
  create.subplot <- function(projections, subplot_title, data) {
    plot <- projections %>%
      pivot_wider(names_from = quantile) %>%
      plot_ly(x = ~date, width = 800, height = plot.height) %>%
      add_ribbons(ymin = ~`0.025`, ymax = ~`0.975`, color = I("lightblue2"), alpha = 0.25,
                hovertemplate = paste0("%{x}: %{y:", y.format, "}<extra>Upper 95% CrI</extra>")) %>%
      add_lines(y = ~`0.025`, alpha = 0,   # An extra trace just for hover text
                hovertemplate = paste0("%{x}: %{y:", y.format, "}<extra>Lower 95% CrI</extra>")) %>%
      add_lines(y = ~`0.5`, color = I("black"),
                hovertemplate = paste0("%{x}: %{y:", y.format, "}<extra>Median</extra>")) %>%
      layout(shapes = lines, annotations = title(subplot_title), showlegend = FALSE,
             hovermode = "x unified")
    if (is.null(data)) return(plot)
    plot %>%
        add_markers(
          data = data,
          x = ~date,
          y = ~True,
          color = I("red"),
          hovertemplate = "%{x}: %{y:.3s}<extra>Actual report</extra>"
        )
  }
  plots <- list("England" = create.subplot(Eng_projection, "England", Eng_data))
  for(subplot in plot.names) {
    if (is.null(data)) {
      plot.data <- NULL
    } else {
      plot.data <- data %>%
        filter(!!sym(by) == subplot) %>%
        group_by(date) %>%
        summarise(True = sum(value))
    }
    plots[[subplot]] <- projections %>%
      filter(by == subplot) %>%
      create.subplot(subplot, plot.data)
  }
  return(subplot(plots, nrows = num.rows))
}
```

## Summary

Real-time tracking of a pandemic, as data accumulate over time, is an essential component of a public health response to a new outbreak. This document reports the work of a joint Public Health England (PHE)---University of Cambridge modelling group to nowcast and forecast COVID-19 infections and deaths, together with estimation of relevant epidemiological quantities for England (by NHS region) and Scotland. These estimates have provided the bases of forecasts supplied to the Scientific Pandemic Influenza sub-group on Modelling (SPI-M) and to regional PHE teams.

## Data Sources

We have been using data from the PHE-generated line-listing, as published on `r format(ymd(date.data), "%d %b")`. 
This consists of a combination of deaths notified to:

* the Demographics Batch Service (DBS), a mechanism that allows PHE to submit a file of patient information to the National Health Service spine for tracing against the personal demographics service (PDS). PHE submit a line list of patients diagnoses with COVID-19 to DBS daily. The file is returned with a death flag and date of death updated (started 20<sup>th</sup> March, 2020).
* NHS England, who report data from NHS trusts relating to patients who have died after admission to hospital or within emergency department settings.
* Health Protection Teams (HPTs), resulting from a select survey created by PHE to capture deaths occurring outside of hospital settings, e.g. care homes (started 23<sup>rd</sup> March, 2020).

This covers all deaths where the fatality had a confirmed positive test. Deaths which have COVID-19 on the death certificate are included in ONS figures but not here; we are looking into their inclusion.


We have separately analysed the data by:

* date of death and by date of report (see Methods section)
* based on daily updates
* stratified into eight age groups: <1, 1-4, 5-14, 15-24, 25-44, 45-64, 65-74, 75+.
* stratified into NHS England regions (North East and Yorkshire, North West, Midlands, East of England, London, South East, South West) and Scotland.

## Key values {.tabset}

### Current R

Value of $R_e$, the average number of secondary infections due to a typical infection today.

```{r Rt-table}
overall_Rt %>%
  add.quantiles(NULL, QUANTILES) %>%
  mutate(region = "England") %>%
  bind_rows(
    get.aggregated.quantiles(Rt, "region", QUANTILES)
  ) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  select(Region = region, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) %>%
  datatable() %>%
  formatRound(2:4)

```

### Number of infections

```{r cum-inf-sum}
format_num <- function(x) {
  return(format(signif(x, 3), big.mark = ",", scientific = FALSE, width = 0))
}

tbl_cum_inf_region <-
  get.aggregated.quantiles(cum_infections, "region", c(0.025, 0.5, 0.975)) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data))
  
tbl_ar_Eng <- get.aggregated.quantiles(cum_infections, NULL, c(0.025, 0.5, 0.975))  %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  mutate(
    population = sum(population$value),
    region = "England"
  )

tbl_inf_region <-
  bind_rows(
    get.aggregated.quantiles(infections, NULL, c(0.025, 0.5, 0.975))  %>%
    pivot_wider(names_from = quantile) %>%
    filter(date == ymd(date.data)) %>%
    mutate(
      region = "England"
    ),
    get.aggregated.quantiles(infections, "region", c(0.025, 0.5, 0.975)) %>%
    pivot_wider(names_from = quantile) %>%
    filter(date == ymd(date.data))
  )%>%
  mutate_at(vars(`0.025`, `0.5`, `0.975`), format_num) %>%
  transmute(
    Region = region,
    Daily_infections = paste0(`0.5`, " (", `0.025`, "--", `0.975`, ")")
  )

```

```{r cum-inf-disp}
bind_rows(
  select(tbl_ar_Eng, -population),
  tbl_cum_inf_region
) %>%
  mutate_at(vars(`0.025`, `0.5`, `0.975`), format_num) %>%
  transmute(
    Region = region,
    Cumulative_infections = paste0(`0.5`, " (", `0.025`, "--", `0.975`, ")")
  ) %>%
  left_join(tbl_inf_region, by = "Region") %>%
  datatable()
```

### Attack rate {.tabset}

The percentage of a given group that has been infected.

#### By region

```{r attack-region}
bind_rows(
  tbl_ar_Eng,
  population %>%
    group_by(region) %>%
    summarise(population = sum(value)) %>%
    left_join(tbl_cum_inf_region, by = c('region' = 'region'))
  ) %>%
  select(Region = region, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`, population) %>%
  mutate(
    Median = Median / population,
    `95% CrI (lower)` = `95% CrI (lower)` / population,
    `95% CrI (upper)` = `95% CrI (upper)` / population
    ) %>%
  select(-population) %>%
  datatable() %>%
  formatPercentage(2:4)
```


#### By age

```{r attack-age}
tbl_cum_inf_age <-
  get.aggregated.quantiles(cum_infections, "age", c(0.025, 0.5, 0.975)) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  select(age, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) 

population %>%
  group_by(age) %>%
  summarise(population = sum(value)) %>%
  left_join(tbl_cum_inf_age, by = c('age' = 'age')) %>%
  mutate(
    Median = Median / population,
    `95% CrI (lower)` = `95% CrI (lower)` / population,
    `95% CrI (upper)` = `95% CrI (upper)` / population
    ) %>%
  select(-population) %>%
  arrange(match(age, age.labs)) %>%
  datatable() %>%
  formatPercentage(2:4)
```
<!--
```{r attack-region-age}
tbl_cum_inf_region <-
  get.aggregated.quantiles(cum_infections, "region", c(0.025, 0.5, 0.975)) %>%
  pivot_wider(names_from = quantile) %>%
  filter(date == ymd(date.data)) %>%
  select(Region = region, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) 
population %>%
  group_by(region) %>%
  summarise(population = sum(value)) %>%
  left_join(tbl_cum_inf_region, by = c('region' = 'Region')) %>%
  #mutate_at(vars(-c(region, population)), function(x, pop) {x / pop}, population)
  mutate(
    Median = 100 * Median / population,
    `95% CrI (lower)` = 100 * `95% CrI (lower)` / population,
    `95% CrI (upper)` = 100 * `95% CrI (upper)` / population
    ) %>%
  select(-population) %>%
  knitr::kable()
```
-->

### IFR


```{r ifr}
today_cum_inf_by_age <- apply(infections[,ymd(date.data)-start.date,,], c("age", "iteration"), sum)
merge.youngest.age.groups <- function(mat, num.to.group = 2) {
  add.function <- function(x) {
    to.merge <- x[1:num.to.group]
    to.preserve <- x[(num.to.group+1):length(x)]
    return(c(sum(to.merge), to.preserve))
  }
  result <- apply(mat, "iteration", add.function)
  return(result)
}

overall_ifr <- today_cum_inf_by_age %>%
  merge.youngest.age.groups() %>%
  t() %>%
  `*`(params$prop_case_to_hosp[parameter.to.outputs,]) %>%
  rowSums() %>%
  `/`(colSums(today_cum_inf_by_age)) %>%
  quantile(QUANTILES) %>%
  enframe(name = "quantile") %>%
  mutate(age = "Overall")

params$prop_case_to_hosp %>%
  as.matrix() %>%
  `colnames<-`(dimnames(deaths)$age) %>%
  as_tibble() %>%
  pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
  group_by(age) %>%
  summarise(value = list(quantile(ifr, QUANTILES))) %>%
  mutate(quantile = list(names(value[[1]]))) %>%
  unnest(c(quantile, value)) %>%
  bind_rows(overall_ifr) %>%
  mutate(
    quantile = parse.percentage(quantile),
    value = paste0(signif(100 * value, 2), "%")
  ) %>%
  pivot_wider(names_from = quantile, values_from = value) %>%
  select(Age = age, Median = `0.5`, `95% CrI (lower)` = `0.025`, `95% CrI (upper)` = `0.975`) %>%
  arrange(match(Age, c("Overall", dimnames(deaths)$age))) %>%
  datatable()

# (params$prop_case_to_hosp %>%
#   as.matrix() %>%
#   `colnames<-`(dimnames(deaths)$age) %>%
#   as_tibble() %>%
#   pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
#   mutate(age = factor(age, levels = dimnames(deaths)$age)) %>%
#   ggplot(aes(x = age, y = ifr)) +
#   geom_violin()) %>%
#   ggplotly()
#   
# params$prop_case_to_hosp %>%
#   as.matrix() %>%
#   `colnames<-`(dimnames(deaths)$age) %>%
#   as_tibble() %>%
#   pivot_longer(everything(), names_to = 'age', values_to = 'ifr') %>%
#   group_by(age) %>%
#   summarise(value = list(quantile(ifr, c(0.025, 0.25, 0.5, 0.75, 0.975)))) %>%
#   mutate(quantile = list(names(value[[1]]))) %>%
#   unnest(c(quantile, value)) %>%
#   mutate(
#     age = factor(age, levels = dimnames(deaths)$age),
#     quantile = parse.percentage(quantile)
#   ) %>%
#   pivot_wider(names_from = quantile, values_from = value, names_prefix = "q") %>%
#   ggplot(aes(x = age, lower=q0.25, upper=q0.75, middle=q0.5, ymin=q0.025, 
#                     ymax=q0.975)) +
#   geom_boxplot(stat = "identity") +
#   scale_y_continuous()
```

### Change in infections incidence {.tabset}

#### Growth rates

```{r growth-rate-func}
today.int <- as.integer(ymd(date.data))
wanted.days <- as.character(c(today.int - 1, today.int, today.int + 1))
calc.growth.rate <- function(data) {
  data %>%
    aperm(c("date", "region", "iteration", "age")) %>%
    `[`(wanted.days,,,) %>%
    apply(c("region", "iteration", "date"), sum) %>%
    as.tbl_cube(met_name = "value") %>%
    as_tibble() %>%
    mutate(
      date = lubridate::as_date(date)
    ) %>%
    arrange(date) %>%
    group_by(region, iteration) %>%
    summarise(growth = ((lead(value) - lag(value)) / (2 * value))[2]) %>%
    group_by(region) %>%
    summarise(Median = median(growth),
              `95% CrI (lower)` = quantile(growth, 0.025),
              `95% CrI (upper)` = quantile(growth, 0.975),
              ) %>%
    mutate(Region = str_replace_all(region, "_", " ")) %>%
    select(Region, everything()) %>%
    select(-region)
}
infections.growth.rates <- calc.growth.rate(infections)
deaths.growth.rates <- calc.growth.rate(deaths)
```

NB: negative growth rates are decline rates. Values are daily changes.

`r knitr::kable(infections.growth.rates, digits = 2, align = 'lccc')`

#### Halving times

Halving times in days, if a region shows growth than value will be Inf.

```{r inf-halving}

infections.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. > 0, Inf, -1/log2(1+.))
  ) %>%
  knitr::kable(digits = 2, align = 'lccc')

```

### Change in deaths incidence {.tabset}


#### Growth rates

NB: negative growth rates are decline rates. Values are daily changes.

`r knitr::kable(deaths.growth.rates, digits = 2, align = 'lccc')`

#### Halving times

Halving times in days, if a region shows growth than value will be Inf.

```{r death-halving}

deaths.growth.rates %>%
  mutate_at(
    vars(-'Region'),
    ~ifelse(. > 0, Inf, -1/log2(1+.))
  ) %>%
  knitr::kable(digits = 2, align = 'lccc')

```


## Infections and deaths {.tabset}

Blue line is date of lockdown (23 Mar), the red line is the date of results (`r format(ymd(date.data), "%d %b")`).

### Infection incidence {.tabset}

#### By region
```{r}
make.plots(infections, by = "region")
```

#### By age
```{r}
make.plots(infections, by = "age")
```

### Cumulative infections {.tabset}

#### By region
```{r}
make.plots(cum_infections, by = "region")
```

#### By age
```{r}
make.plots(cum_infections, by = "age")
```

### Deaths incidence {.tabset}

#### By region
```{r}
make.plots(noisy_deaths, by = "region", data = dth.dat)
```

#### By age
```{r}
make.plots(noisy_deaths, by = "age", data = dth.dat)
```

### Cumulative deaths {.tabset}

#### By region
```{r}
make.plots(cum_deaths, by = "region")
```

#### By age
```{r}
make.plots(cum_deaths, by = "age")
```

### Rt
```{r}
get.combined.Rt <- function(...) overall_Rt
make.plots(Rt, by = "region", y.format = ".2f", combine.to.England = get.combined.Rt)
```

<!--
## Methods {.tabset}

### Overview

Model figure here

### Transmission

SEIR model parameterisation

### Contact matrices

How we account for the lockdown

### Disease Severity

What proportion of infections will lead to death

### Disease Reporting

When will we know about the deaths, relative to disease onset.


-->

