---
title: "Snapshot Analysis"
author: "Paul Birrell, Joshua Blake, Edwin van Leeuwen, Joel Kandiah, MRC Biostatistics Unit COVID-19 Working Group, Daniela De Angelis."
date:   "`r Sys.time()`"
output: 
    html_document:
        code_folding: hide
---

Analysis of a snapshot of the results from the Real-time Nowcasting and Forecasting model with input data until `r str.date` and the model state on `r str.date`.

```{r, include = FALSE}
knitr::opts_chunk$set(
    fig.width = 12,
    fig.height = 8,
    warning = FALSE,
    error = FALSE
)
    
# Load in necessary libraries
library(tidyverse)
library(lubridate)

# setwd("~/Snapshot Analysis/results_20220218/Prev655SeroNHSBT_All_NHScutoff_IFR6bp_11wk2_prev14-0PHE_matrices2_20220218_timeuse_household_admissions_no_deaths_chain2")
theme <- theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text = element_text(size = 14),
        title = element_text(size = 18, face = "bold"),
        strip.text = element_text(size = 14),
        strip.placement = "outside"
      )

age.levels <- c(
  "<1yr",
  "1-4",
  "5-14",
  "15-24",
  "25-44",
  "45-64",
  "65-74",
  "75+"
)

state <- state |> 
  filter(age.labs != "<1yr") |> 
  mutate(age.labs = factor(age.labs, levels = age.levels, labels = age.levels, ordered = T)) |> 
  filter(state.names != "plambda")

dark_back <- T

if(dark_back) {
  bg_col <- "gray20"
} else {
  bg_col <- "white"
}

if(region.type == "ONS") {
  region_abbrev <- c(
    "East Midlands" = "EM",
    "East of England" = "EE",
    "London" = "GL",
    "North East" = "NE",
    "North West" = "NW",
    "South East" = "SE",
    "South West" = "SW",
    "West Midlands" = "WM",
    "Yorkshire and The Humber" = "YH"
  )
}


```

## Plots of Proportions of Individuals in each Class

These plots estimate the proportion of individuals in each class using the total population as the number of individuals in the primary grouping(s), age and/or region, across all vaccination statuses.

- The Latent infection class represents individuals in the E classes in the epidemic model
- The Prevalent infection class represents individuals in the I classes and R+ class in the epidemic model

## {.tabset}

### By Age

```{r}
state.levels <- c(
  "Susceptible, Never Infected" = "Fully susceptible",
  "Susceptible, previously infected" = "Susceptible, previously infected",
  "Latent Infection" = "Latent infection",
  "Prevalent Infection" = "Prevalent infection",
  "Infection-Acquired Immunity" = "Infection-acquired immunity"
  )

with.doses <- c(
  paste0(" with ", c(0,"1-2",3,4), " vaccine doses")
)

## state_colours <- c(  "Fully susceptible" = "#3cb44b",
##   "Susceptible, previously infected" = "#ffe119",
##   "Latent infection" = "#4363d8",
##   "Prevalent infection" = "#f58231",
##   "Infection-acquired immunity" = "#fabed4"
## )

combined_levels <- crossing(
  state_levels = names(state.levels),
  with_doses = with.doses
) |> 
  mutate(levels = paste0(state_levels, with_doses)) |> 
  mutate(state_levels = factor(state_levels, levels = names(state.levels), ordered = T),
         with_doses = factor(with_doses, levels = with.doses, ordered = T)
         ) |> 
  arrange(desc(state_levels), desc(with_doses))

state_colours <- c(
"#226912",
"#5c9549",
"#92c480",
"#caf4b9",
"#ccb414",
"#ddc550",
"#edd67d",
"#fbe8a7",
"#1e2ea6",
"#7764c3",
"#b4a0e1",
"#ede1ff",
"#e67a2e",
"#f09c5c",
"#f8bc8a",
"#ffdbbc",
"#7a3858",
"#a56a88",
"#d19eba",
"#ffd4ee"
)

dose_aggregation <- function(dose){
  idx <- which(dose %in% 1:2)
  dose[idx] <- "1-2"
  dose
}

names(state_colours) <- combined_levels$levels

summary_simple_state_ages <- state |> 
  filter(!is.na(state.names), state.names != "plambda") |> 
  mutate(state_text_old = state.text, dose.numbers = dose_aggregation(dose.numbers), dose.numbers_old = dose.numbers, state.text = factor(state.text, levels = state.levels, labels = names(state.levels), ordered = T)) |> 
  unite(col = "state.text", state.text, dose.numbers, sep = " with ") |> 
  mutate(state.text = paste0(state.text, " vaccine doses")) |> 
  mutate(state.text = factor(state.text, levels = combined_levels$levels, ordered = T)) |>
  group_by(age.labs, region, iteration, state.names, state.text, state_text_old, dose.numbers_old, popn) |>
  summarise(value = sum(value)) |>
  group_by(age.labs, iteration, state.names, state.text, state_text_old, dose.numbers_old) |> 
  summarise(value = sum(value), popn = sum(popn)) |> 
  group_by(age.labs, popn, state.text, iteration, state_text_old, dose.numbers_old) |> 
  summarise(value = sum(value)) |> 
  group_by(age.labs, popn, state.text, state_text_old, dose.numbers_old) |> 
  summarise(value = mean(value)) |>  
  ungroup()

p1 <- summary_simple_state_ages |> 
  ggplot(aes(x = age.labs, y = value / popn * 100, colour = state.text, text = paste0("State: ", state.text, "\nAge: ", age.labs, "\nProportion of Population: ", scales::label_number(accuracy = 0.01)(value / popn * 100), "%"))) +
  scale_colour_manual(values = state_colours) +
  geom_col(aes(fill = after_scale(colour)))  +
  theme +
  labs(x = "Age", y = "Proportion of\nPopulation (%)", colour = "State") +
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 15, face = "bold"),
        panel.background = element_rect(fill = bg_col),
          panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank())

# p1 + theme(axis.title.y=element_text(angle=0))


plotly::ggplotly(p1, tooltip = "text") |> 
  plotly::layout(legend = list(font = list(family = "Open Sans", size = 15, color = "black")),
                  legend_title = list(font = list(family = "Open Sans", size = 14, color = "black")))

```

### By Region

```{r}
summary_simple_state_ages <- state |> 
  filter(!is.na(state.names), state.names != "plambda") |> 
  mutate(state_text_old = state.text, dose.numbers = dose_aggregation(dose.numbers), dose.numbers_old = dose.numbers, state.text = factor(state.text, levels = state.levels, labels = names(state.levels), ordered = T)) |> 
  unite(col = "state.text", state.text, dose.numbers, sep = " with ") |> 
  mutate(state.text = paste0(state.text, " vaccine doses")) |> 
  mutate(state.text = factor(state.text, levels = combined_levels$levels, ordered = T)) |>
  group_by(age.labs, region, iteration, state.names, state.text, state_text_old, dose.numbers_old, popn) |>
  summarise(value = sum(value)) |>
  mutate(region = str_replace_all(region, "_", " ")) |> 
  group_by(region, iteration, state.names, state.text) |> 
  summarise(value = sum(value), popn = sum(popn)) |> 
  group_by(region, popn, state.text, iteration) |> 
  summarise(value = sum(value)) |> 
  group_by(region, popn, state.text) |> 
  summarise(value = mean(value)) |>  
  ungroup()

p1 <- summary_simple_state_ages |> 
  mutate(region_labelled = factor(region, levels = names(region_abbrev), labels = region_abbrev, ordered = T)) |> 
  ggplot(aes(x = region_labelled, y = value / popn * 100, colour = state.text, text = paste0("State: ", state.text, "\nRegion: ", region, "\nProportion of Population: ", scales::label_number(accuracy = 0.01)(value / popn * 100), "%"))) +
  scale_colour_manual(values = state_colours) +
  geom_col(aes(fill = after_scale(colour)))  +
  theme +
  labs(x = "Region", y = "Proportion of\nPopulation (%)", colour = "State") +
  # theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 15, face = "bold"),
        panel.background = element_rect(fill = bg_col),
          panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank())
  # theme(axis.title.y=element_text(angle=0))

# p1 + theme(axis.title.y=element_text(angle=0))


plotly::ggplotly(p1, tooltip = "text")|> 
  plotly::layout(legend = list(font = list(family = "Open Sans", size = 15, color = "black")),
                  legend_title = list(font = list(family = "Open Sans", size = 14, color = "black")))

```

### By Age and Region

```{r, fig.height=16}
summary_simple_state_ages <- state |> 
  filter(!is.na(state.names), state.names != "plambda") |> 
  mutate(state_text_old = state.text, dose.numbers = dose_aggregation(dose.numbers), dose.numbers_old = dose.numbers, state.text = factor(state.text, levels = state.levels, labels = names(state.levels), ordered = T)) |> 
  unite(col = "state.text", state.text, dose.numbers, sep = " with ") |> 
  mutate(state.text = paste0(state.text, " vaccine doses")) |> 
  mutate(state.text = factor(state.text, levels = combined_levels$levels, ordered = T)) |>
  group_by(age.labs, region, iteration, state.names, state.text, state_text_old, dose.numbers_old, popn) |>
  summarise(value = sum(value)) |>
  mutate(region = str_replace_all(region, "_", " ")) |> 
  group_by(age.labs, region, iteration, state.names, state.text) |> 
  summarise(value = sum(value), popn = sum(popn)) |> 
  group_by(age.labs, region, popn, state.text, iteration) |> 
  summarise(value = sum(value)) |> 
  group_by(age.labs, region, popn, state.text) |> 
  summarise(value = mean(value)) |>  
  ungroup()

p1 <- summary_simple_state_ages |> 
  ggplot(aes(x = age.labs, y = value / popn * 100, colour = state.text, text = paste0("State: ", state.text, "\nRegion: ", region, "\nAge: ", age.labs, "\nProportion of Population: ", scales::label_number(accuracy = 0.01)(value / popn * 100), "%"))) +
  scale_colour_manual(values = state_colours) +
  geom_col(aes(fill = after_scale(colour)))  +
  theme +
  labs(x = "Age", y = "Proportion of\nPopulation (%)", colour = "State") +
  facet_wrap(~region, ncol = 2) +
  # theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 15, face = "bold"),
        panel.background = element_rect(fill = bg_col),
          panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank()) +
  theme(panel.spacing.y = unit(1, "line"))

# p1 + theme(axis.title.y=element_text(angle=0))


plotly::ggplotly(p1, tooltip = "text") |> 
  plotly::layout(legend = list(font = list(family = "Open Sans", size = 15, color = "black")),
                  legend_title = list(font = list(family = "Open Sans", size = 14, color = "black")))
```
